### 2014-03-31 paper/main.tex

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{fullpage}

\begin{document}
\author{Julian Bangert and Nickolai Zeldovich}
\date{}
\title{Nail: A practical interface generator for binary formats}

\maketitle

\input{abstract}
\input{background}
\input{design}
\input{impl}
\input{eval}
\input{future}

\bibliography{main,locasto,login,phrack,short,langsec}
\bibliographystyle{plain}

\end{document}