### 2014-03-31 paper/main.tex
some typesetting tweaks

\documentclass[10pt,twocolumn]{article}
\usepackage[10pt,nocopyright]{sigmin}
\usepackage[letterpaper, breaklinks, pdfborder={0 0 0}]{hyperref}
\usepackage[square,comma,numbers,sort&compress]{natbib}
\usepackage{times}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{fullpage}

\renewcommand{\ttdefault}{pxtt}   %% better \tt font

%% Ensure ligatures (e.g., ``fine official flag'') can be copy/pasted from PDF.
\input{glyphtounicode}
\pdfgentounicode=1

\begin{document}
\author{Julian Bangert and Nickolai Zeldovich}
\date{}
\title{Nail: A practical interface generator for binary formats}

\maketitle

\input{abstract}
\input{background}
\input{design}
\input{impl}
\input{eval}
\input{future}

\bibliography{main,locasto,login,phrack,short,langsec}
\bibliographystyle{plain}

\end{document}