### 2014-04-01 paper/concl.tex
stub conclusion

\section{Conclusion}

This paper presented the initial design and implementation of
\textit{Nail}, an XXX...