### 2014-04-02 paper/design.tex
Design

\section{Design}
\label{s:design}

\paragraph{Overview}

To alleviate these problems, Nail provides a richer set of combinators that not
only describes the grammar of the external protocol, but also induce an internal
model.

A central design decision of Nail is that there is a semantic bijection between
the model exposed to the programmer and the bytes input and output up to
syntactic equivalence for unambiguous grammars. More precisely, the parser is
the generators inverse, so parsing the generators output will yield the
generators input, but generating the parsers output does not necessarily yield
the parsers input. To understand why this makes sense, consider a grammar for a
text language that tolerates white space or a binary protocol that tolerates
arbitrarily long padding\footnote{Say, the physical layer of most communication
  protocols is a possibly infinite sequence of symbols that are syntactically
  nil followed by a pre-determined synchronisation sequence and the actual
  contents of the transmission.}. In that case, the program semantics should be
independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded
fields \emph{constants}. Currently, Nail always makes a default choice when
there are multiple options to express a constant, however Nail could be extended
to allow a grammar-specific plug-in to make these choices, say for faster
alignment, streaming applications when data is not ready or visual appearance in
human-readable protocols. Similarly, Nail does not preserve the layout of
objects referred to by their offsets. If the grammar contains no constants and
offset fields, there is a proper isomorphism between model and protocol.

Nail evaluates the combinators and produces type declarations for the internal
model, the \textit{parser}, a function to parse a sequence of bytes into an
instance of the model and the \textit{generator}, a function to create a
sequence of bytes from an instance of the model. In the following, we present
Nail's combinators.

\paragraph{Fundamental parsers}
The elementary parsers of Nail are the same as those of Hammer, signed and
unsigned integers with arbitrary lengths up to 64 bits. Note that is possible to
define parsers for sub-byte lengths, e.g. to parse the 4-bit data offset within
the TCP header. Integer parsers return their value in the
smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
Examples: \texttt{uint4}, \texttt{int32}

Integer parsers can be constrained to fall either within an (inclusive) range of
values or be an
element of a set of allowed values. \\
Examples:\\
\texttt{uint8 | 1..3} (8 bit unsigned integer $1\leq x \leq 3$) ,\\
\texttt{uint16| ..512} (unsigned 16-bit integer $\leq 512$),\\
\texttt{int32 | [1,255,512]} (signed 32 bit integer with value 1, 255 or 512).
Invalid values or prematurely reaching the end of input raise an error when
parsing input or generating output and abort the procedure.

\paragraph{Sequence: Structure}
Nails fundamental concept is the structure combinator. It contains a list of
named parsers and unnamed constant parsers  . The parser invokes each field 
in sequence and returns a structure containing the result of each parser.

\begin{tabular}{ll}
  \textbf{Nail} & \textbf{Data Model}\\
  \begin{minipage}{1.2in}
\begin{verbatim}
header = {
 id uint16
 qr uint1
 opcode uint4
 aa uint1 
 tc uint1
 rd uint1
 ra uint1
 uint3 = 0
 rcode uint4
}
\end{verbatim} 
  \end{minipage}
  & 
  \begin{minipage}{1.2in}
\begin{verbatim}
struct header{
    uint16_t id;
    uint8_t qr;
    uint8_t opcode;
    uint8_t aa;
    uint8_t tc;
    uint8_t rd;
    uint8_t ra;
    uint8_t rcode;
};
\end{verbatim} 
  \end{minipage}
  \\
\end{tabular}
\paragraph{Repetition: Many,SepBy}
The many combinator takes a parser and applies it repeatedly until it fails,
returning an array of the inner parsers results. SepBy additionally takes a
constant parser, which it applies in between
parsing two values,but not before parsing the first value or after parsing the last.\\
Examples: \texttt{many uint8}\\ \texttt{sepBy uint8=',' (many uint8| '0'..'9')}
(which recognises comma-separated lists of numbers).

\paragraph{Semantic choice: choose}
We extend PEG's ordered choice combinator with a tag for each choice. The parser
attempts to parse each option in the order they are specified in the field and
stores the result in a tagged union. If an option fails, the parser backtracks
to the beginning of the choice combinator's input. Care must be taken that the
choices do not overlap, because Nail always picks to the first successfully
parsed choice. If two options overlap, generated output for the latter option is
not necessarily understood identically by the parser. However, actual data
formats are usually not ambiguous in this sense.

\begin{verbatim}
choose {
   A = uint8 | 1.. 8
   B = uint16 | ..256
}
\end{verbatim}

\subsection{Constant parsers}
Nail also features \emph{constant parsers}, which are not affecting the internal
representation. Constant parsers can appear instead of structure fields and as
separators in the SepBy combinator.

The simplest constant parser is an integer or array of integers with fixed
value, e.g. \texttt{uint8=0}, \texttt{many uint8=[1,2]}. A convenience notation
for strings is also supported: \texttt{many uint8 = "foo"}.

\paragraph{Wrap combinator.} When implementing real protocols with Nail, we often found that many structures
consist of many constant parsers and only one named field. This pattern is
common in binary protocols which use fixed headers to denote the type of data
structure to be parsed.  In order to keep the internal representation cleaner,
we introduced the wrap combinator, which takes a sequence of parsers containing
exactly one non-constant parser. The parser and generator act as though the wrap
combinator was a sequence of parsers, but the data model does not wrap the
single value in another structure, making the application code more concise.\\
Example: \\ \verb+<uint8='"';many int8|'a'..'z'; uint8='"'>+  parses a quoted
lowercase word into an array - excluding the quotation marks.

\paragraph{Constant combinators.}
In some protocols, there might be many ways to represent the same constant field
and there is no semantic difference between different syntaxes.  Nail therefore
supports choice (\texttt{|| uint8 = 0 || uint16 = 0xFFFF}) and repetition
(\texttt{many (uint8= ' ')}). This is particularly useful for handling padding or whitespace.

As discussed above, choosing to use these combinators on constant parsers
removes the bijection between byte-strings and our data model, as there are
multiple byte-strings that correspond to the same internal data structure and
the generator has to choose one of these representations. As such, constant
combinators are the generator dual of ambiguous choice combinators in the
parser, because they lead to ambiguities in the generator.

\subsection{Dependent fields}
Another problem for parser generators is that binary protocols often contain
length and offset fields. While conventional parsing algorithms can deal with
bounded offset fields - a finite automaton can count a bounded integer, and we
can feed the (finite) input multiple times to the finite automaton. However,
this implementation is both time-inefficient - why feed many bytes into the
automaton that will just be skipped - and very cumbersome to express with
current parser generators. Therefore, if languages with offset fields are parsed
with parser generators, the only currently feasible way is to add 'ad-hoc' hacks
such as changing the input pointer of the generated parser on the fly from
semantic actions.

Nail will properly support both offset and length fields and much of the
following discussion applies to both , although the current prototype only
implements lengths, which we will focus on.

We call length or offset fields \textit{dependent fields}, because during
parsing, another parser depends on them, and while generating output, their
value depends on some other structure in the data model. Dependency fields
appear in a structure combinator as would any other integer field, but their
name begins with an '@' sign. Dependency fields are not exposed in the data
model, but transparently computed.

A dependency field has to appear in the grammar before it can be used. 
\paragraph{Length fields: n_of}
The length combinator takes a dependency field and a parser, evaluating the
parser as often as the value of the dependency field, returning an array of the
values. When generating output, it emits the array and writes its length to the
dependency field.
\paragraph{Offsets: offset} 
The offset combinator takes a dependency field and a parser. It moves the parser
to the position specified in the offset field and invokes the inner parser, and
then moves the input back to its original position.
While generating output, all structures referred to by offset are generated
after the main structure and the dependent offset fields patched up.

\paragraph{Checksums: raw_depend}
In many data formats, some values depend on external representation, such as
checksums and cryptographic signatures. While it would be possible to extend
our constraint language to be powerful enough to support such constructs, we
would essentially be building a separate, Turing complete language that has all
the same pitfalls existing programs have. Therefore, we intend to  allow the programmer to
very carefully escape Nail's programming model and write a function that takes a
pointer to the dependent value and a range of bytes. 
This feature is not implemented in the current prototype. \\
Example:\\
 \verb+{ data many uint8; @checksum uint32+\\
 \verb+raw_depend @checksum data crc32 }+\\
 \verb+/*bool crc32(uint32_t *out, uint8_t *in);*/+
Because this feature compromises Nail's security guarantees, it should only be
used in limited circumstances and with carefully prepared checksum functions. 
\subsection{Example Grammars}
To further familiarise the reader with Nail, we provide two example grammars.
\paragraph{UTF-16}
This grammar recognises valid UTF-16 strings and exposes an array of code
points.

\begin{tabular}{ll}
  \textbf{Nail} & \textbf{Data Model}\\
  \begin{minipage}{3in}
\begin{verbatim}
utfstring = many choose {
         SUPP= {
               lead uint16 | 55296..56319
                          // 0xD800..0xDBFF
               trail uint16 | 56320..57343
                          // 0xDC00..0xDFFF
               }
         BASIC = uint16 | !55296..57343
       }
\end{verbatim}
  \end{minipage}
  & 
  \begin{minipage}{2in}
\begin{verbatim}
struct utfstring {
 struct {
   enum  {SUPP,BASIC} N_type;
   union {
    struct {
     uint16_t lead;
     uint16_t trail;
    } SUPP;
    uint16_t BASIC;
   };
 }*elem;
 size_t count;
};
\end{verbatim} 
  \end{minipage}
  \\
\end{tabular}

\paragraph{DNS packet}
This grammar recognises DNS packets without label compression, as per RFC1035
\begin{verbatim}
labels = <many {@length  uint8 | 1..255 
                label n_of @length uint8 }
             uint8 = 0>
question= {   labels labels
              qtype uint16 | 1..16 
              qclass uint16 | [1,255]
               }
answer  = {
             labels labels
             rtype uint16 | 1..16
             class uint16 | [1]
             ttl uint32
             @rlength uint16 
             rdata n_of @rlength uint8 
             }
dnspacket = {id uint16
              qr uint1
              opcode uint4
              aa uint1 
              tc uint1
              rd uint1
              ra uint1
              uint3 = 0
              rcode uint4
              @questioncount uint16
              @answercount uint16
              authoritycount uint16   // Ignored
              additionalcount uint16  //Ignored       
              questions n_of @questioncount question         
              responses n_of @answercount answer
            }
\end{verbatim}