### 2014-04-02 paper/concl.tex
Conclusion, eval, future work

\section{Conclusion}
\label{s:concl}
This paper presented the initial design and implementation of \textit{Nail}, a usable interface
generator for data formats. Nail allows programmers to avoid memory corruption and ambiguity
vulnerabilities while reducing effort in parsing and generating real world protocols and file
formats. We have demonstrated Nail's expressive power by implementing DNS and Ethernet.

Our code is open source on GitHub\footnote{\url{https://github.com/jbangert/nail/}}.