### 2014-04-02 paper/concl.tex
some relwk nits; one XXX regarding hammer vs nail

\section{Conclusion}
\label{s:concl}

This paper presented the initial design and implementation of
\textit{Nail}, an XXX...

We hope that by reducing programmer effort required to use a parser
generator, Nail will enable more application developers to use generated
parsers in practice.