### 2014-04-02 paper/concl.tex
conclusion

\section{Conclusion}
\label{s:concl}

This paper presented the initial design and implementation of
\textit{Nail}, a parser generator that aims to minimize the amount
of programmer effort.  Nail achieves this by reducing the expressive
power of the grammar, maintaining a \emph{semantic bijection} between
data formats and internal representations, and allowing programmers to
specify \emph{structural dependencies} on the data format.  This avoids
the need for programmers to write explicit semantic actions for input
and output processing.  We hope that by reducing programmer effort
required to use a parser generator, Nail will enable more application
developers to use generated parsers in practice.  Our prototype source
code is available at \url{https://github.com/jbangert/nail}.