### 2014-04-09 paper/osdi14/concl.tex
copy langsec paper to osdi14

\section{Conclusion}
\label{s:concl}

This paper presented the initial design and implementation of
\textit{Nail}, an interface generator for data formats.  Nail helps
programmers to avoid memory corruption and ambiguity vulnerabilities
while reducing effort in parsing and generating real-world protocols
and file formats.  Nail achieves this by reducing the expressive
power of the grammar, maintaining a \emph{semantic bijection} between
data formats and internal representations, and allowing programmers to
specify \emph{structural dependencies} in the data format.  Preliminary
experience with implementing a DNS server using Nail suggests that this
is a promising approach.

The source code for our prototype of Nail is available at
\url{https://github.com/jbangert/nail}.