### 2014-04-23 paper/osdi14/abstract.tex
Suggestion for better abstract

\begin{abstract}

  We present \textit{Nail}, an interface generator that allows programmers to safely parse and
  generate protocols defined by a Parser-Expression based grammar. Nail uses a richer set of parser
  combinators that induce an internal representation, obviating the need to write semantic actions.
  Nail also introduces \textit{dependency fields} and \textit{stream transformations} that capture
  protocol features such as size and offset fields, check-sums and compressed data that are hard to
  express in existing protocol languages.

\end{abstract}