### 2014-05-01 paper/osdi14/concl.tex
done with a pass

\section{Conclusion}
\label{s:concl}

This paper presented the design and implementation of \textit{Nail},
an interface generator for data formats.  Nail helps programmers avoid
memory corruption and ambiguity vulnerabilities while reducing effort
in parsing and generating real-world protocols and file formats.
Nail achieves this by reducing the expressive power of the grammar,
maintaining a \emph{semantic bijection} between data formats and internal
representations.  Nail captures complex data formats by introducing
\emph{dependent fields}, \emph{streams}, and \emph{transforms}.
Using these techniques, Nail is able to support DNS packet and ZIP file
formats, and enables applications to handle these data formats in many
fewer lines of code.

% The source code for our prototype of Nail is available at
% \url{https://github.com/jbangert/nail}.