### 2014-05-01 paper/osdi14/eval.tex

\section{Evaluation}
\label{s:eval}

\input{fig-ethernet}

In our evaluation of Nail, we answer four questions:

\begin{itemize}

\item Can Nail grammars support real-world data formats, and
      are Nail's techniques critical to handling these formats?

\item How much programmer effort is required to build an
      application that uses Nail for data input and output?

\item Does using Nail for handling input and output improve
      application security?

\item Does Nail achieve acceptable performance?

\end{itemize}

\subsection{Data formats}
\label{s:eval-formats}

To answer the first question, we used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
of code for their Nail grammars, and the challenging features that make
the protocols difficult to parse with state-of-the-art parser generators.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.
In the rest of this subsection, we describe the DNS and Zip grammars in
more detail, focusing on how Nail's features enable us to support these
formats.

\begin{figure}[tb]
\centering
\begin{tabular}{lrl}
\toprule
\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
\midrule
DNS packets & 48+64 & Label compression,\\
  & & count fields \\
ZIP archives & 92+78 & Checksums, offsets, \\ 
  & & variable length trailer, \\
  & & compression \\
Ethernet  & 16+0\phantom{0} & --- \\
ARP       & 10+0\phantom{0} & --- \\
IP        & 25+0\phantom{0} & Total length field, options \\
UDP       &  7+0\phantom{0} & Checksum, length field \\
ICMP      &  5+0\phantom{0} & Checksum \\
\bottomrule
\end{tabular}

\caption{Protocols, sizes of their Nail grammars, and challenging aspects
of the protocol that cannot be expressed in existing grammar languages.
A + symbol counts lines of Nail grammar code (before the +) and lines of
C code for protocol-specific transforms (after the +).}
\label{fig:eval-protocols}
\end{figure}

\paragraph{DNS.}

Figure~\ref{fig:dns-xform} presents the Nail grammar for DNS packets.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.

One challenging aspect of DNS packets lies in the count fields (\cc{qc},
\cc{ac}, \cc{ns}, and \cc{ar}), which represent the number of questions
or answers in another part of the packet.  Nail's \cc{n_of} combinator
handles this situation easily, which would have been difficult to handle
for other parsers.

Another challenging aspect of DNS is label
compression~\cite[\S4.1.4]{RFC:1035}.  Label compression is used to reduce
the size overhead of including each domain name multiple times in a DNS
reply (once in the question section, and at least once in the response
section).  If a domain name suffix is repeated, instead of repeating that
suffix, the DNS packet may write a two-bit marker sequence followed by
a 14-bit offset into the packet, indicating the position of where that
suffix was previously encoded.

\input{fig-dnsfull}

Handling label compression in existing tools, such as Bison or Hammer,
would at best be very awkward, because some ad-hoc trick would have to
be used to re-position the parser's input stream.  Keeping track of the
position of all recognized labels would not be enough, as the offset field
may refer to any byte within the packet, not just the beginning of labels.
For this reason, the DNS server used as the example for Hammer does not
support compression.

\begin{figure}
\smaller[0.5]
\input{code/dns-xform}
\caption{Signatures of stream transform functions for handling DNS label compression.}
\label{fig:dns-xform}
\end{figure}

In contrast, Nail is able to handle label compression, by using a stream
transform; the signatures of the two transform functions are shown
in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
decompresses the DNS label stream by following the offset pointers.
When generating a packet, this transform receives the current suffix as
an input, and scans the packet so far for previous occurrences, which
implements compression.

\paragraph{ZIP files.}

An especially tricky data format is the ZIP compressed archive
format~\cite{pkzip}.  ZIP files are normally parsed end-to-beginning. At
the end of each ZIP file is an \emph{end-of-directory header}. This header
contains a variable-length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid
length field is found. Many ZIP implementations disagree on how to find
this header in confusing situations, such as when the comment contains
the magic number~\cite{wolf:berlinsides-zip}.  This end-of-directory
header contains the offset and size of the \emph{ZIP directory}, which
is an array of \emph{directory entry headers}, one for every file
in the archive.  Each entry stores file metadata, such as file name,
compressed and uncompressed size, and a checksum, in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and the compressed file
contents follow it immediately.

\input{fig-zip}

Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, and memory constraints
made it impossible to keep the ZIP directory in memory or the archive
was split across multiple disks.  However, care must be taken that the
metadata is consistent. For example, vulnerabilities could occur if
the length in the central directory is used to allocate memory and the
length in the local directory is used to extract without checking that
they are equal first, as was the case in the Python ZIP
library~\cite{cve-python-zip-maybe}.

\begin{figure}[tb]
\smaller[0.5]
\input{code/zip-eod-xform}
\caption{Signatures of stream transform functions for handling the
end-to-beginning structure of ZIP files.}
\label{fig:zip-eod-xform}
\end{figure}

Figure~\ref{fig:zip-extract} shows an abbreviated version of our ZIP
file grammar.  The ZIP grammar is a good example of how transformations
capture complicated syntax in a real-world file format; existing parser
languages cannot handle a file format of this complexity.

The \cc{zip\_file} grammar first splits the entire file stream into two
streams based on the \cc{zip\_end\_of\_directory} transform (whose two
C function signatures are shown in Figure~\ref{fig:zip-eod-xform}).
\cc{zip\_end\_of\_directory_parse} finds the end-of-directory header
as described above, by scanning the file backwards, and splits the
file into two streams, one containing the end-of-directory header and
one containing the file contents.  The header stream is then parsed,
according to the \cc{apply} rule, using the \cc{end\_of\_directory}
parser.  The file stream, to which all the offsets in the metadata refer,
is passed as an argument to that parser.  The directory header is then
parsed, and the offset and size transforms provided by Nail are used
to isolate the actual directory from the file stream.  Each directory
entry in turn points to a local file header.

The \cc{file} grammar for a ZIP file entry is particularly interesting.
It receives as arguments the file size, checksum, and compression method
of the file from the directory header.  However, this same information
is duplicated in the file entry, so the grammar uses the Nail-supplied
\cc{depend} transform to check whether the two values are equal. Unlike
most other transforms, \cc{depend} does not consume or produce strings; it
only checks that two dependent fields are equal when parsing, and assigns
the value of the second field to the first when generating.  This ensures
that the programmer does not have to worry about inconsistencies when
handling the internal representation of a ZIP file.

Immediately following the file entry is the compressed data.  Because
most compression algorithms operate on unbounded streams of data, Nail
decompresses data in two steps.  First, it isolates the compressed
data from the rest of the stream by using the \cc{size} transform,
which operates on the current stream, meaning it will consume data
starting at the current position of the parser in the input.  Second,
Nail invokes a custom \cc{zip\_compression} transform that implements
the appropriate compression and decompression functions based on the
specified compression method. These functions are otherwise oblivious
to the layout or metadata of the file.

\subsection{Programmer effort}
\label{s:eval-effort}

%code size unzip 6.0
%2821 lines

To evaluate how much programmer effort is required to build an
application that uses Nail, we implemented two applications---a
DNS server, and an
\cc{unzip} program---based on the above grammars, and compared code
size with comparable applications that process data manually, using
\cc{sloccount}~\cite{sloccount}.  We also compare the code size of our
DNS server to a DNS server written using the Hammer parsing framework,
although it does not fully support DNS (e.g., it lacks label compression,
among other things).  Figure~\ref{fig:effort} summarizes the results.

\begin{figure}[tb]
\centering
\begin{tabular}{@{~}lrr@{~}l@{~}}
\toprule
\textbf{Application}
  & \textbf{LoC w/ Nail}
  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
\midrule
DNS server
  & 295
  & 683
  & (Hammer parser) \\

%  &
%  & 10,000
%  & (DJBDNS) \\

\cc{unzip}
  & 220
  & 1,600
  & (Info-Zip) \\
\bottomrule
\end{tabular}
\caption{Comparison of code size for three applications written in Nail,
         and several existing alternative implementations.}
\label{fig:effort}
\end{figure}

\paragraph{DNS.}

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
only 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.
From this, we conclude that Nail leads to much more compact code for
dealing with DNS packet formats.

% It is hard to compare the programming effort required to implement
% our DNS server to that of a real world DNS server, since we implement
% less functionality: we do not send the additional glue records and
% we did not implement any configuration options.  However, the closest
% in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
% which aims to be a minimalist, highly secure, authoritative-only DNS
% server. The latest release of djbdns, including various support tools,
% is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]

\paragraph{ZIP.}

We implemented a ZIP file extractor in 50 lines of C code, together with
92 lines of Nail grammar and 78 lines of C code implementing two stream
transforms (one for the DEFLATE compression algorithm with the help of
the zlib library, and one for finding the end-of-directory header).

Because more recent versions of ZIP have added more features, such
as large file support and encryption, the closest existing tool
in functionality is the historic version 5.4 of the Info-Zip unzip
utility~\cite{infozip} that is shipped with most Linux distributions.
The entire unzip distribution is about 46,000 lines of code, which is
mostly optimized implementations of various compression algorithms and
other configuration and portability code.  However, unzip isolates the
equivalent of our Nail tool in the file \cc{extract.c}, which parses the
ZIP metadata and calls various decompression routines in other files. This
file measures over 1,600 lines of C, which suggests that Nail is highly
effective at reducing manual input parsing code, even for the complex
ZIP file format.

\subsection{Security}

We use a twofold approach to evaluate the security of applications
implemented with Nail. First, we analyze a list of CVE's related to the
ZIP file format and argue how our ZIP tools based on Nail are immune
against those vulnerability classes. Second, we present the results of
fuzz-testing our DNS server.

\paragraph{ZIP analysis.}

The ZIP format has been associated with many vulnerabilities, and the
PROTOS Genome project found numerous security vulnerabilities in most
implementations of ZIP and other archive formats that are directly
related to input handling.

Between May 2011 and May 2014, the CVE database~\cite{cve-database} lists
83 vulnerabilities when searching for the string ``ZIP''. Of these, 42
were completely unrelated to the archive format, for example reporting
vulnerabilities revealing ZIP codes of other users. 26 vulnerabilities
were related to ZIP archives, but not to input processing, such as
trusting arbitrary update ZIP files sent to a server or vulnerabilities
that can be accessed through manipulating files in a ZIP container. This
includes several vulnerabilities allowing another files to replaced while
they are extracted from ZIP files because insufficiently secure temporary
directories are used. Of the 15 remaining vulnerabilities pertaining
to ZIP files, 14 vulnerabilities were related to input processing. The
other vulnerability, which falls outside of Nail's purview, concerns a
ZIP packer choosing a weak legacy cryptosystem even when configured to
use AES.

Nail parsers are immune by design to vulnerabilities like these 14.
Ten of those vulnerabilities are memory corruption attacks, that could
also be avoided by using safer languages.  Most of the buffer overflows
seem to result from inconsistent offset fields being manipulated
without checking.  Because Nail's generated code checks offsets before
reading and does not expose any untrusted pointers to the application,
it protects against memory corruption attacks by design.

The remaining four vulnerabilities do not involve memory corruption,
but rather parser ambiguities.  Two of them relate to implementations
mis-parsing ZIP files, particularly in virus scanners. One other
vulnerability is in a Python ZIP library, which uses both the size
field in the file entry and in the header, without checking for their
consistency.  This results in a denial-of-service through endless loops
or incorrectly extracted ZIP files.

A similar attack was used by the fourth vulnerability, the infamous
Android master key bug~\cite{saurik-masterkey} that completely
bypassed Android security through parser inconsistencies between the
ZIP handler that checks signatures for privileged applications and the
ZIP implementations that ultimately extract those files. Thus, valid
application bundles could be modified to include malicious applications
without breaking their signatures. Instead of both components containing
an ad-hoc archive implementation, a single Nail grammar would be reusable
across both implementations.

\paragraph{DNS fuzzing.}

To provide additional assurance that Nail parsers are free of memory
corruption attacks, we ran the DNS fuzzer provided with the Metasploit
framework~\cite{mspframework} on our DNS server, which sent randomly
corrupted DNS queries to our server for 4 hours, during which it did
not crash or trigger the stack or heap corruption detector.

\subsection{Performance}

To evaluate whether Nail-based parsers are compatible with good
performance, we compare the performance of our DNS server to that of ISC
BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
We simulate a load resembling that of an authoritative name server.
First, we generate domain names consisting of one or two labels randomly
selected from an English dictionary, and one label that is one of
three popular top-level domains (\cc{com}, \cc{net}, and \cc{org}).
Second, we randomly selected 90\% of these domains and created a zone
file that mapped these domain names to \cc{127.0.0.1}.  Finally, we used
the \cc{queryperf} tool provided with BIND to query each domain between
zero and three times, using a DNS server running on the local machine.
We used a single core of an Intel i7-3610QM system with 12GB of RAM.
The benchmark tool kept at most 20 queries outstanding at once, and
was configured to repeat the same randomized sequence of queries for
one minute.  We repeated each test seven times with 50,000 domain names,
restarting each daemon in between; we also repeated the tests with 1
million domain names, and found similar results.  We also performed one
initial dry run to warm the file system cache for the zone file.

The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
our Nail-based DNS server can achieve higher performance and lower latency
than BIND\@.  Although BIND is a much more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server, we believe our results demonstrate that Nail's parsers are not
a barrier to achieving acceptable performance.

\input{fig-perf-dns}