### 2014-05-01 paper/osdi14/future.tex

\section{Future work}
\label{s:future}

One problem with the current design of Nail is that the design of the
grammar dictates the internal structure of the software.  This makes
changing grammars or adding Nail to existing software awkward.
One possible solution to this problem would be to implement a concept
similar to relational lenses~\cite{bohannon2006relational}, which would
allow the data model to be ``seen'' by the rest of the program through
an isomorphism.  Such an isomorphism would still be much more concise
than two sets of semantic actions, while allowing changes in syntax,
alternative representations, and adaption to legacy systems.

% Finally, we would like to demonstrate the capabilities of Nail by
% implementing various binary formats ``notorious'' for their insecurity
% in Nail. Nail was designed with the idioms of formats such as PDF and
% PNG in mind. Ultimately, we want to provide examples of successful Nail
% parsers throughout a network stack, from a user-space TCP stack to a
% PNG de-compressor.