### 2014-05-02 paper/osdi14/fig-motivation.tex
footnote moved

\begin{figure*}
\centering
\begin{tabular}{lllr}
\toprule
 \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
\midrule
  Memory corruption & 2013-5660 & Buffer overflow & 11\\
  Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
  Semantic misunderstanding & 2014-2319 & Weak cryptography used even if user selects AES & 1\\
  \multicolumn{3}{l}{\bf Total vulnerabilities related to .zip processing }& \bf 16  \\
\bottomrule

\end{tabular}
\caption{Classification of known vulnerabilities in the CVE database
between 2010 and 2014 related to the search term ``ZIP'' and involving
the ZIP file format.}
\label{fig:motivation} 
\end{figure*}