### 2014-09-17 paper/osdi14/fig-motivation.tex

\begin{figure*}
\centering
\smaller[0.5]
\input{table-motivation.tex}
\caption{Classification of vulnerabilities in the CVE database
from 2010 to May 2014 containing the term ``ZIP'' and involving
the ZIP file format.}
\label{fig:motivation} 
\end{figure*}