### 2014-11-14 paper/login/concl.tex

\section{Conclusion}
\label{s:concl}

 We presented the design and implementation of \textit{Nail},
a tool for parsing and generating complex data formats based on a
precise grammar. This helps programmers avoid
memory corruption and inconsistency vulnerabilities while reducing effort
in parsing and generating real-world protocols and file formats.
 Nail captures complex data formats by introducing
\emph{dependent fields}, \emph{streams}, and \emph{transforms}.
Using these techniques, Nail is able to support DNS packet and ZIP file
formats, and enables applications to handle these data formats in many
fewer lines of code.

Nail and all of the applications and grammars developed in this paper
are released as open-source software, available at
\url{https://github.com/jbangert/nail}. A more detailed discussion of our design and our results is
available in~\cite{bangert:osdi-nail14}.