### 2014-11-14 paper/login/table-motivation.tex

\begin{tabular}{lllr}
\toprule
 \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
\midrule
  Memory corruption & CVE-2013-5660 & Buffer overflow & 11\\
  Parsing inconsistency & CVE-2013-1462 & Multiple virus scanners interpret ZIP files incorrectly & 4\\
  Semantic misunderstanding & CVE-2014-2319 & Weak cryptography used even if user selects AES & 1\\
\midrule
  \multicolumn{3}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
\bottomrule

\end{tabular}