### 2014-11-25 paper/osdi14/poster.tex

\documentclass[25pt, a0paper, portrait, margin=0mm, innermargin=15mm,
blockverticalspace=10mm, colspace=15mm, subcolspace=8mm]{tikzposter}

\usepackage{times}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{color}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{relsize}
\usepackage{fancyvrb}
\usepackage{pgfplots}
\usepackage{adjustbox}
\title{Nail: A practical tool for parsing and generating data formats}
\author{Julian Bangert and Nickolai Zeldovich }
\institute{MIT CSAIL}
%\titlegraphic{Logo}
\usetheme{Envelope}

\input{code/fmt}

\begin{document}
\maketitle
\begin{columns}
\column{.5}

\block{Parsing vulnerabilities}{
\begin{itemize}
\item Hand-written parsers introduce bugs.
\item Different parsers not equivalent
\begin{itemize}
\item Evasi0n jailbreaks on iOS.
\item PKI layer cake.
\item Android master key.
\end{itemize}
\end{itemize}

\innerblock{ZIP: A tricky file format}{
  \includegraphics[width=0.43\columnwidth]{ZipHorizontal.png}
}

\innerblock{Case study of ZIP vulnerabilities in the CVE database:}{

\begin{tabular}{lllr}
\toprule
 \bf Classification  & \bf Example description & \bf Count\\
\midrule
  Memory corruption  & Buffer overflow & 11\\
  Parsing inconsistency & Virus scanners interpret ZIP files incorrectly & 4\\
  Semantic misunderstanding & Weak cryptography used even if user selects AES & 1\\
\midrule
  \multicolumn{2}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
\bottomrule

\end{tabular}
}
}

\block{Existing parsers}{   
Automatically generated parsers, such as  \texttt{bison} or Hammer are
\begin{itemize}
\item Immune to classes of bugs (e.g. memory corruption)\
\item Easier to use
\item Re-useable, eliminating parser ambiguities
\end{itemize} 
However, traditional parsers are inconvenient to use, because the programmer has to write all the
yellow-colored parts:

\includegraphics{StateOfTheArt.png}

}
\block{Nail}{
Nail grammars describe both the format and a data type to represent it, an idea introduced by data
description languages such as PacketTypes.

\includegraphics{Nail.png}

\innerblock{Dependent Fields}{
\begin{itemize}
\item Redundant Information in protocol
\begin{itemize}
\item Length field
\item Repeated information
\end{itemize}
\item Confuses application, e.g.: 
\begin{itemize}
\item Android master key: Uses local name length for extracting, directory for signature
\item PKI layer cake: Mix implicit (terminator) string length and length field.
\end{itemize}
\end{itemize}
Dependent Fields \textbf{hide redundant information}and automatically
verify and generate it.
}
\innerblock{Streams and Transformations}{
\begin{itemize}
 \item Existing parsers are linear, consuming input front to back.
\item  Nail grammars feature multiple streams.
\item \emph{Transformations} modify and create streams.
\item Pair of functions operating on streams and dependent fields.
\item Standard library of Transformations enough for many formats.
\item Programmer can write their own for complex formats.
\end{itemize}
}
}
% \block{Design Goals}{
%   \begin{itemize}
%     \item Reduce programmer effort
%   \begin{itemize}
%       \item Semantic combinators describes data type and format.
%       \item Generate output. Semantic bijection, preserves meaning, but can change representation.
%   \end{itemize}
%      \item Handle complex formats
%   \begin{itemize}
%        \item Dependency Fields: hide redundant and structural information.
%        \item Transforms: Extensibility mechanism to capture many features without complicating the
%          grammar language.
%        \end{itemize}
%   \end{itemize}
%}
% \block{Semantic Bijection}{
%   Traditional bijections don't make sense for some data formats. For example, offsets are and should
%   be discarded during parsing. Nail parsers and generators therefore form a looser 'semantic
%   bijection', in which parser(generator(x)) = x, but not necessarily vice versa.

% }
\column{.5}
\block{Nail Grammar for .ZIP }{
\innerblock{Nail Grammar}{
\small
\input{code/zip.nail}
}
\innerblock{Excerpt of the generated API}{
\input{code/zip-out}
}
}
\block{Evaluation}{
\begin{tabular}{lllllll}
\toprule
\textbf{Protocol}&\textbf{Grammar} & \textbf{Transform}& \textbf{Application} & \textbf{App LoC} &
\textbf{Total LoC } &
\textbf{Alternative}\\
\midrule
DNS & 48&64 & Server & 186 & 298 & 683 (Hammer) \\
\multicolumn{3}{l}{} & Resolver &  97 & 209 & >3000 (libc) \\
ZIP & 92&78 & Extractor & 49 & 219 & 1600(Info-ZIP)\\
Ethernet & 16 & 0 & \multicolumn{4}{l}{}\\
ARP & 10 & 0 & \multicolumn{4}{l}{}\\
IP & 25 & 0 & \multicolumn{4}{l}{}\\
UDP & 7 & 0 & \multicolumn{4}{l}{}\\
ICMP & 5 & 0 & \multicolumn{4}{l}{}\\

\end{tabular}

Nail and our example applications are available at

\url{https://github.com/jbangert/nail}.
}

\end{columns}

\end{document}