### 2015-08-10 paper/thesis/fig-stream-api.tex

\begin{figure*}
  \begin{tabular}{lp{2.75in}}
    \toprule
    \bf Method & \bf Description \\
    \midrule
    \texttt{bool check(uint n)} & Check if the current position is at least $n$ bits before the end of the stream. \\
    \texttt{uint read\_bits(uint n)} &  Read $n$ bits from the stream. Check($n$) must be true. Does not move the current position.\\
    \texttt{uint read\_bits\_little(uint n)} &  Like read\_bits, but interprets the value as little endian.\\
    \texttt{void advance(uint n)} & Advances the current position by $n$ bits.\\
    \texttt{void backup(uint n))} & Reverses the current position by $n$ bits.\\
    \texttt{pos\_t get\_pos(uint n))} & Returns the current position.\\
    \texttt{void reposition(pos\_t old))} & Reverts the current position to a value formerly returned from get\_pos().\\

    \bottomrule
    \end{tabular}
\label{fig:stream-api}
\caption{The operations Nail parsers invoke on their input streams.}
\end{figure*}